---
title: "*Emotionality in a second language: It's a matter of time* (Opitz & Degner, 2012): an adaptation"
subtitle: "Data Science for Linguists: Final Project"
author: "Eva María Corregidor Luna"
institute: "Department of Spanish and Portuguese, Rutgers University"
date: "2025/05/05 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: ["hygge", "rutgers", "rutgers-fonts"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r include=FALSE}
library(tidyverse)
library(ggplot2)
library(knitr)
library(here)
library(ggthemes)
```

```{r include=FALSE}
data_raw <- read.csv(here("data", "raw", "response_latency_data.csv"))
```

```{r include=FALSE}
data_tidy <- data_raw |>
  separate(
    col = id,
    into = c("id", "language"),
    sep = "_") |>
  separate (
    col = item,
    into = c("word_type", "number"),
    sep = "_") |>
  rename(accuracy = response) |>
  select(id, language, word_type, accuracy, latency)
```

#*Emotionality in a second language: It's a matter of time* (Opitz & Degner, 2012)

- Different sense of emotionality in a person’s first (L1) and second language (L2).
- **Participants**: 33 bilinguals (FR-GE).
- **Task**: Lexical Monitoring Task (Yes/No Decision).
- **EEG Recording**: Data recorded while processing pleasant, unpleasant and neutral words in L1 and L2. Differences in processing recorded through the early posterior negativity (EPN).
- **Results**: the affective valence of L2 words is processed in a less immediate way due to delayed lexical access; interference in a highly integrated L1/ L2 mental lexicon."
---
#An adaptation for ds4ling
##Research Questions & Hypothesis

**RQ1**: Are emotion concepts in the L2 processed later and less accurately than in the L1?

**H1**: Accuracy in L1 will be greater than accuracy in L2; latency will be greater when performing the task in the L2 than in the L1.
---
#An adaptation for ds4ling
##Participants
- 150 Spanish-English bilinguals (75 women).

- Aged 25–35 years.

- Recruited among the administrative personnel from Rutgers University.

- Right-handed, normal or corrected-to-normal vision, reported no history of neurological or language disorders.

- **Late L2 learners (AoA = 12)**.

- By the time of testing...
    - **Highly proficient in Spanish.**
    - Mean of self-rated frequency of **weekly L2 use of 65%**.
    - At least one long stay (>6 months) in a L2 country.
---
#An adaptation for ds4ling
##Tasks & Stimuli
**1. Language History Questionnaire**

Used to obtain information about the bilingual profile, L2 acquisition, L2 use and L2 frequency. It contained 7 multiple-choice questions, 8 questions with numerical scales and 5 open questions (total of 20 questions).
---
#An adaptation for ds4ling
##Tasks & Stimuli
**2. Lexical Monitoring Task**

For the ENGLISH version of the task...
- Stimuli were obtained from an English words database (Imbault et al., 2020) were words were rated by participants based on valence and arousal. 
- From the database, the **50 words** with the highest valence level were chosen to be the **emotional words**. Similarly, the **50 words** with the lowest valance level were chosen to be the **non-emotional words**. 
- All words were nouns and were matched in word length, frequency of use. For the **pseudo-words**, **50 words** were created using Artificial Intelligence (AI).
---
#An adaptation for ds4ling
##Tasks & Stimuli
**2. Lexical Monitoring Task**
For the SPANISH version of the task...
- Words were translated by 5 native speakers and those translations that matched were used.

The task was divided in five rounds, each with 30 trials (10 emotion words, 10 non-emotion words, 10 pseudo-words).
---
#An adaptation for ds4ling
##EEG Recording

Early emotion processing is captured at the word level by the Early Posterior Negativity (EPN)...
- Elicited around 200-300 ms post-stimulus.
- Mapped in the occipital cortex.
- An index of rapid emotional attention (Luck, 2014).
- Here, use to capture response-accuracy and latency (and amplitude?).
---
#An adaptation for ds4ling
##Statistical Analysis

**Dependent Variables**
- *Accuracy* in L1 AND L2 (right/wrong)
- *Latency* (ms, 280-430)

**Predictors**
- *Language* (l1, l2)
- *Word type* (emotion, non-emotion, and pseudo-words [only in the task, not for analysis])

**To keep in mind...**
- Accuracy is a binary outcome, a Yes/No decision AND a count vector.
- Latency does not range from -∞ to ∞.
- Predictors (language, word-type) are both categorical variables.
---
#An adaptation for ds4ling
##Statistical Analysis: Accuracy Model
A generalized linear model + binary regression

*Accuracy* ~ *Task Language* + *Word Type*
```{r echo=TRUE}
mod_accuracy <- glm(
  accuracy ~ language + word_type, 
  data = data_tidy, 
  family = "binomial"
)
```
---
#An adaptation for ds4ling
##Statistical Analysis: Accuracy Model
```{r echo=FALSE}
summary(mod_accuracy)
```
---
```{r echo=FALSE, warning=FALSE}
#| dpi: 400
#| fig-align: center
#| max-width: "75%"
#| out-width: "75%"
data_tidy |>
  ggplot() +
  aes(x = accuracy, y = language, color = word_type) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "glm", method.args = list(family = "binary")) +
  scale_fill_viridis_d(begin = 0.3, end = 0.8) +
  scale_color_viridis_d(begin = 0.3, end = 0.8) +
  ds4ling::ds4ling_bw_theme()
```
---
#An adaptation for ds4ling
##Statistical Analysis: Latency Model
A general linear model (spoiler: plots are more fun)

*Latency* ~ *Task Language* + *Word Type*
```{r echo=TRUE}
mod_latency <- lm(
  accuracy ~ language + word_type, 
  data = data_tidy)
```
---
#An adaptation for ds4ling
##Statistical Analysis: Latency Model
```{r echo=FALSE}
summary(mod_latency)
```
---
```{r echo=FALSE}
#| out-width: "85%"
#| max-width: "85%"
#| dpi: 400
#| fig.align: "center"
data_tidy |>
  ggplot(aes(x = language, y = latency, color = word_type)) +
  geom_jitter(height = 0, width = 0.2, alpha = 0.4) +
    stat_summary(
  fun.data = mean_sdl,
  geom = 'pointrange'
) +
  facet_grid(. ~ word_type) +
  scale_fill_viridis_d(begin = 0.3, end = 0.8) +
  scale_color_viridis_d(begin = 0.3, end = 0.8) +
  ds4ling::ds4ling_bw_theme()
```
---
#An adaptation for ds4ling
##Considerations & Current Challenges
- Complexity of the model: **need** to understand it.
- Filter for pseudowords.
- Plots.
- Adding an slope?
  - For **accuracy**: *Accuracy* ~ *Task Language* x *Word Type* (+ 1 | item)
  - For **latency**: *Latency* ~ *Task Language* x *Word Type* (+ 1 | item)
- **Extra layers**: proficiency, amplitude.
---
#An adaptation for ds4ling
Thank you for listening! Please use this time to provie any **suggestion** to improve the project and to ask any **questions** you might still have. 🙂
